Legacy Data Confounds Modern Studies

### 2018-04-12 Draft3.tex
two columns

\documentclass[12pt]{amsart}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{epstopdf}
\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
\AppendGraphicsExtensions{.tiff}

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\lat}[1]{\textcolor{blue}{LAT: #1}}
\newcommand{\todo}[1]{\textcolor{blue}{todo: #1}}

@@

\maketitle
\tableofcontents
\clearpage
			\section{Introduction}
	\subsection{Larger Sample Sizes}			
The last 5 years has seen an increase in the number of individuals genotyped through private companies or medical research cohorts. 
Larger sample sizes allow researchers to identify finer resolution statistically significant differences between groups of individuals. 
Mutations associated to a disease can be identified by comparing the genomes of healthy individuals to those afflicted by a disease. 
The mutations commonly found in patients and rarely in controls might be associated to the disease in question.
However, demonstrating that these mutations are biologically relevant can be difficult.
Especially with the increasing size of cohorts, spurious associations are increasingly becoming an issue. 
For this reason, careful consideration must be taken when including individuals from different ancestral origins in these association studies.
Benign mutations at high frequencies in one populations might be exceedingly rare in another population.
Therefore, population wide differences in mutations must be included as covariates to avoid spurious associations.

@@

\begin{figure}
\includegraphics[width=\hsize,keepaspectratio]{MapQualOverTime.jpg}

\caption{A) The average mapping quality of each individual per population included in the 1000 Genomes sequencing project. The average mapping quality per individual was obtained from the raw alignment index data for each sample. The X axis is ranked by populations with the lease to the most variance, followed by average mapping quality per individual. B) Same data as in A) except the X axis is now sorted by sequencing date, and the colors indicate the sequencing centers that produced the data for each individual.}
\label{Figure1}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\hsize,keepaspectratio]{GWAS_Qual_GenomeWide_JPT.jpg}
\caption{Genome wide association of the average quality of mapped bases for the Japanese individuals included in the 1000 Genomes Project. This GWAS included 104 individuals, \todo{000} of them were sequenced in phase 1 while the rest were sequenced in phase 3 of the 1000 Genomes Project. The same analysis was performed independently for each of the populations in the 1000 Genomes Project. }
 \label{Figure2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\hsize,keepaspectratio]{SNPOverlap6.jpg}

\caption{Overlap of SNPs identified independently to be associated with quality. Not surprisingly, the populations that have the most low-quality individuals also have the most low-quality variants. What is interesting here, is that the same variants identified as being low quality independently in each population are found in other populations. }
  \label{Figure3}
\end{figure}

>>>>

\documentclass[12pt,twocolumn]{article}
\usepackage[margin=2.2cm]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

%\geometry{a4paper} % or letter or a5paper or ... etc

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\lat}[1]{\textcolor{blue}{LAT: #1}}
\newcommand{\todo}[1]{\textcolor{blue}{*#1*}}

@@

\onecolumn
\maketitle
\tableofcontents
\clearpage
\twocolumn
			\section{Introduction}
	\subsection{Larger Sample Sizes}			
The last 5 years has seen an increase in the number of individuals genotyped through private companies or medical research cohorts. 
Larger sample sizes allow researchers to identify finer resolution statistically significant differences between groups of individuals. 
Mutations associated to a disease can be identified by comparing the genomes of healthy individuals to those afflicted by a disease. 
The mutations commonly found in patients and rarely in controls might be associated to the disease in question.
However, demonstrating that these mutations are biologically relevant can be difficult.
Especially with the increasing size of cohorts, spurious associations are increasingly becoming an issue. 
For this reason, careful consideration must be taken when including individuals from different ancestral origins in these association studies.
Benign mutations at high frequencies in one populations might be exceedingly rare in another population.
Therefore, population wide differences in mutations must be included as covariates to avoid spurious associations.

@@

\begin{figure*}
\includegraphics[width=\hsize,keepaspectratio]{MapQualOverTime.jpg}

\caption{A) The average mapping quality of each individual per population included in the 1000 Genomes sequencing project. The average mapping quality per individual was obtained from the raw alignment index data for each sample. The X axis is ranked by populations with the lease to the most variance, followed by average mapping quality per individual. B) Same data as in A) except the X axis is now sorted by sequencing date, and the colors indicate the sequencing centers that produced the data for each individual.}
\label{Figure1}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\hsize,keepaspectratio]{GWAS_Qual_GenomeWide_JPT.jpg}
\caption{Genome wide association of the average quality of mapped bases for the Japanese individuals included in the 1000 Genomes Project. This GWAS included 104 individuals, \todo{000} of them were sequenced in phase 1 while the rest were sequenced in phase 3 of the 1000 Genomes Project. The same analysis was performed independently for each of the populations in the 1000 Genomes Project. }
 \label{Figure2}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\hsize,keepaspectratio]{SNPOverlap6.jpg}

\caption{Overlap of SNPs identified independently to be associated with quality. Not surprisingly, the populations that have the most low-quality individuals also have the most low-quality variants. What is interesting here, is that the same variants identified as being low quality independently in each population are found in other populations. }
  \label{Figure3}
\end{figure*}
